\plaintitle

### 2016-05-20 Paper/sp2-paper.tex
Update paper

\documentclass{sigchi}

@@

% Load basic packages
\usepackage{balance}  % to better equalize the last page
\usepackage{graphics} % for EPS, load graphicx instead 
\usepackage[T1]{fontenc}
\usepackage{txfonts}
\usepackage{mathptmx}
\usepackage[pdftex]{hyperref}
\usepackage{color}
\usepackage{booktabs}
\usepackage{textcomp}
% Some optional stuff you might like/need.
\usepackage{microtype} % Improved Tracking and Kerning
% \usepackage[all]{hypcap}  % Fixes bug in hyperref caption linking
\usepackage{ccicons}  % Cite your images correctly!
\usepackage[utf8]{inputenc} % for a UTF8 editor only

@@

% Paper metadata (use plain text, for PDF inclusion and later
% re-using, if desired).  Use \emtpyauthor when submitting for review
% so you remain anonymous.
\def\plaintitle{University of the Philippines Los Baños Online Student Information System}
\def\plainauthor{Jan Keith Darunday, Ferriel Lisandro Melarpis, Toni-Jan Keith Monserrat, and Reginald Neil Recario}
\def\emptyauthor{}
\def\plainkeywords{Authors' choice; of terms; separated; by
  semicolons; include commas, within terms only; required.}
\def\plaingeneralterms{Documentation, Standardization}

@@

%Jan Keith Darunday, Ferriel Lisandro Melarpis, Toni-Jan Keith Monserrat, and Reginald Neil Recario
\numberofauthors{4}
\author{%
  \alignauthor{Jan Keith Darunday\\
    \affaddr{Los Ba\~{n}os, Philippines}\\
    \email{jcdarunday@up.edu.ph}}\\
  \alignauthor{Ferriel Lisandro Melarpis\\
    \affaddr{Los Ba\~{n}os, Philippines}\\
    \email{fbmelarpis@up.edu.ph}}\\
  \alignauthor{Toni-Jan Keith Monserrat\\
    \affaddr{Los Ba\~{n}os, Philippines}\\
    \email{tonijanmonserrat@gmail.com}}\\
%   \alignauthor{Reginald Neil Recario\\
%     \affaddr{Los Ba\~{n}os, Philippines}\\
%     \email{rcrecario@up.edu.ph}}\\
}

@@

In campuses under the University of the Philippines (UP) System, students have 
more freedom to choose the subjects that they want to enroll in compared to other
universities. UP employs the Revitalized General Education Program (RGEP) that
requires students to take General Education (GE) courses. They have the ability to
choose which of these courses they would like to enlist in.
These GE courses are courses that allow students from any degree program to enroll
unlike major courses that are only available to certain degree programs. The
ability to choose one’s subjects implied that almost all students will need to
manually fill-up their list of courses to enroll to before the beginning of 
every semester. Moreover, courses are divided into sections each of which have 
a limited number of slots for students that it can accommodate. This meant that 
not only will the students need to manually check for conflicts in their 
schedules, they will also need to check if the course that they want to enlist 
in still has slots remaining.

The complexity of this process along with the increasing availability of Internet access to
the public incited the development of computerized Online Registration Systems (ORS). These
systems allowed the enlistment, cancellation, and swapping of slots for courses over the Internet.
These systems also did the checking of conflicting subjects and availability of slots automatically
so that students will not need to do them manually.
In the University of the Philippines Los Baños (UPLB), the Online Registration System that
is used as of the writing of this paper is called SystemOne. This system has undergone various
revisions and improvements as old technology become deprecated by new ones and problems
encountered in previous versions are fixed. The most recent one identifies itself as the third
version and is code-named “Decaf”. However, this system did not prove to be efficient enough
for the student population of UPLB despite several rewrites and revisions.

As a solution, we propose that there needs to be a complete rewrite of 
the system. Because SystemOne is originally written in the PHP programming 
language which has been known to be slow, using a faster programming language 
would be an improvement. SystemOne also has its database connection as its 
bottleneck but it tries to solve this by using memcache as a caching system. 
However, this wasn't efficient enough as queries executed by the backend is too
dynamic which makes the results hard to cache. In order to alleviate the effects
of this bottlneck, we propose that three differnt database/storage systems will
be used. One will be a RAM-based storage system with high-speed read/write that
does not need to support queries, another would be a database system that 
supports fast queries and uses a non-volatile memory storage, the last one would
be a full-fledged relational database system that will be used for persistent
storage.

@@

The system will utilized three database systems. These database systems are as
follows:
\begin{enumerate}
\item High-Speed Read and Write Storage System - This will be shared across all
clusters. This needs to high speed because it will used to store information
that change very fast and is read repeatedly such as the number of slots for
every subject.

\item High-Speed Read and Medium-Speed Write Relational Database System - Also 
shared across all clusters, this will contain data and information that are not
changed often but are read a repeatedly a lot such as the information for subjects
and their dependencies, user accounts, and recommended courses.

\item Persistent Relational Database System - This will be used as the permanent
storage for all data. The other two storage and database systems will synchronize
to this database at specific intervals.
\end{enumerate}

@@

The backend will be a server that serves a Representational State Transfer (REST)
Application Programming Interface (API) through the Hypertext Transfer Protocol
(HTTP). The system's backend shall have multithreading support. A thread will be 
spawned for every successful HTTP connection. This thread will handle its own
associated connection and wait for a command before processing it and giving a 
reply. If a command is not received or completed after a specified amount of
time, a timeout will occur and the connection will be dropped along with its
corresponding thread being deallocated from the memory.

@@

Every time a modification or addition is done on a student's schedule, the 
backend will check for conflicts before executing the specified action. 
To do this, a bitwise AND will be done on the current user's schedule and the
schedule of the subject to be inserted. If the resulting value is not 0 then
an error reply will be sent.

@@

The system will be designed to support HTTP Requests from specific domains
where the frontend code will be stored through the Access-Control-Allow-Origin header. 
This is to be able to outsource the load added by frontend data being served as 
it will take a significant amount of networking bandwidth to serve the images and
scripts included. This will also allow the frontend to have heavier images and 
scripts without affecting the performance of the main server.

@@

The frontend will be a static web application that fires REST requests to the
backend server. The Javascript Object Notation (JSON) will be used to specify parameters
for commands that require sending non-anonymous data while a normal GET request
will be used for reading anonymous data. The frontend will also have an enlistment queue
where the user can setup which subjects he/she would like to enlist and then
send them as a batch to the server. This would lessen the number of connections
that the server will receive lessening the number of threads running in the 
backend.

The design for the frontend will try to follow Google's Material Design 
Specification although not strictly. Along with this, it shall support mobile
platforms by implementing Responsive Design.

@@

The system shall be evaluated by its stability, efficiency, ease of use, and
integration.

Its stability will be evaluated by running a fuzzer software on the different
commands that the backend server accepts. If the fuzzer finds strings that 
causes the backend to crash, show an error, or produce the wrong output then the
system fails the stability test. If this happens, the errors shall be fixed and
the system will again undergo the stability evaluation.

The system's efficiency will be evaluated by running benchmarks on the system.
Information such as requests per second, time per request, and transfer rate
of each function will be collected. The system will also be tested if can 
handle medium and heavy workloads by using load testing tools and services.
These tests will use combinations of the different functions at ratios expected
during actual use of the system.

The system's ease of use will be evaluated through a survey along with additional
criteria about other aspects of the system.
The survey will be done with at least 500 respondents that will test the system
and rate it under the following criteria:
\begin{enumerate}
 \item the ability of the user-interface to be understandable without a manual;
 \item the way how buttons in the user-interface are positioned;
 \item the animations embedded into the user-interface
 \item the responsiveness of the system; and
 \item the additional features of the system.
\end{enumerate}

\section{Timeline}
\begin{center}
    \begin{tabular}{ | l | p{5cm} |}
    \hline
      Date & Activity \\ \hline
      Day 0-7 & Planning of System Framework and Database Structures \\ \hline
      Day 8-14 & Implementation of Basic Backend Functions Without Conflict Checking \\ \hline
      Day 15-30 & Implementation of Conflict Checking and Integration with Backend \\ \hline
      Day 31-45 & Development of Static Web Application Interface \\ \hline
      Day 46-60 & Implementation of Advanced Features \\ \hline
      Day 61-68 & Stress Testing and Fuzzing \\ \hline
      Day 69-76 & Adjustments from the Results of Stress Testing and Fuzzing \\ \hline
      Day 77-81 & Alpha Test of Test Users \\ \hline
      Day 82-96 & Adjustments from the Results of Alpha Test \\ \hline
      Day 97-103 & Beta Release \\ \hline
    \end{tabular}
\end{center}

% REFERENCES FORMAT
% References must be the same font size as other body text.
\bibliographystyle{sigchi}
\bibliography{paper}

>>>>

\documentclass[draft]{sigchi}

@@

% Load basic packages
\usepackage{balance}  % to better equalize the last page
\usepackage{graphics} % for EPS, load graphicx instead 
\usepackage[T1]{fontenc}
\usepackage{placeins}
\usepackage{txfonts}
\usepackage{mathptmx}
\usepackage[pdftex]{hyperref}
\usepackage{color}
\usepackage{booktabs}
\usepackage{textcomp}
\usepackage{pgfplots}
% Some optional stuff you might like/need.
\usepackage{microtype} % Improved Tracking and Kerning
% \usepackage[all]{hypcap}  % Fixes bug in hyperref caption linking
\usepackage{ccicons}  % Cite your images correctly!
\usepackage[utf8]{inputenc} % for a UTF8 editor only

@@

% Paper metadata (use plain text, for PDF inclusion and later
% re-using, if desired).  Use \emtpyauthor when submitting for review
% so you remain anonymous.
\def\plaintitle{Flash Crowd Online Student Enrollment System}
\def\plainauthor{Jan Keith Darunday, Ferriel Lisandro Melarpis, Toni-Jan Keith Monserrat, and Reginald Neil Recario}
\def\emptyauthor{}
\def\plainkeywords{Authors' choice; of terms; separated; by
  semicolons; include commas, within terms only; required.}
\def\plaingeneralterms{Documentation, Standardization}

@@

%Jan Keith Darunday, Ferriel Lisandro Melarpis, Toni-Jan Keith Monserrat, and Reginald Neil Recario
\numberofauthors{4}
\author{%
  \alignauthor{Jan Keith Darunday\\
    \affaddr{Los Ba\~{n}os, Philippines}\\
    \email{jcdarunday@up.edu.ph}}\\
  \alignauthor{Ferriel Lisandro Melarpis\\
    \affaddr{Los Ba\~{n}os, Philippines}\\
    \email{fbmelarpis@up.edu.ph}}\\
  \alignauthor{Toni-Jan Keith Monserrat\\
    \affaddr{Los Ba\~{n}os, Philippines}\\
    \email{tonijanmonserrat@gmail.com}}\\
  \alignauthor{Reginald Neil Recario\\
    \affaddr{Los Ba\~{n}os, Philippines}\\
    \email{rcrecario@up.edu.ph}}\\
}

@@

In academic institutions with large student populations such as campuses under
the University of the Philippines (UP) System, the implementation of an online registration
system for students is considered to be one of the biggest challenges as the system needs
to be capable of handling the surge of network connections from the whole student population
during the online registration period. Some of these systems were created by students studying 
in their respective universities and some of the others were created by software development
companies paid by the university. However, many of these systems failed
to address need of the institution that it serves due to issues like lack of budget, lack of development time,
use of wrong or outdated technologies, and lack of consultation of what the users actually need.

This research paper will use the enrollment system and online registration system in
the University of the Philippines as a context.

In the UP System, students are free to enlist into General Education (GE)
courses. These GE courses are courses that allow students from any degree program to enlist into
unlike major courses that are only available to certain degree programs. The
ability to choose one’s subjects implied that almost all students will need to
manually fill-up their list of courses to enroll to before the beginning of 
every semester. Moreover, courses are divided into sections each of which have 
a limited number of slots for students that it can accommodate. This meant that 
not only will the students need to manually check for conflicts in their 
schedules, they will also need to check if the course that they want to enlist 
in still has slots remaining.

The complexity of this process along with the increasing availability of Internet access to
the public incited the development of computerized Online Registration Systems (ORS). These
systems allowed the enlistment, cancellation, and swapping of slots for courses over the Internet.
These systems also did the checking of conflicting subjects and availability of slots automatically
so that students will not need to do them manually.
In the University of the Philippines Los Baños (UPLB), the Online Registration System that
is used as of the writing of this paper is called SystemOne. This system has undergone various
revisions and improvements as old technology become deprecated by new ones and problems
encountered in previous versions are fixed. The most recent one identifies itself as the third
version and is code-named “Decaf”. However, this system did not prove to be efficient enough
for the student population of UPLB despite several rewrites and revisions.
This paper seeks to improve open this kind of online registration by fixing the problems encountered
by SystemOne as a basis.

Because SystemOne is originally written in the PHP programming 
language which has been known to be slow, using a faster programming language 
would be an improvement. SystemOne also has its database connection as its 
bottleneck but it tries to solve this by using memcache as a caching system. 
However, this wasn't efficient enough as queries executed by the backend is too
dynamic which makes the results hard to cache. There are also lots of workarounds
that try to solve the race conditions that occur. Our system will try to solve this
by choosing a more feasible database system for sudden surges of requests.

@@

As an information system, the speed of the system depends heavily on the speed
of the database management system since most of the requests that it handles
do modification of data. Because of this, Redis was chosen as it is a database
management system that has low runtime overhead and very fast request-serving
speeds with benchmarks showing hundreds of thousands of requests per second for
atomic functions.

In order to achieve its speed, Redis sacrifices data integrity by writing
and reading all data to and from the main memory of the system. This leads to possible
data loss upon the occurance of a power interruption while Redis hasn't yet synchronized
its data to permanent storage. However, this should not matter as long as the server
has proper uninterruptible power supply systems in place. Even if such an event occurs,
Redis uses sequential writing which will simply cancel out the latest writes causing a
rollback rather than corrupting the whole database.

@@

The backend is a server that serves a Representational State Transfer (REST)
Application Programming Interface (API) through the Hypertext Transfer Protocol
(HTTP). The system's backend shall have multithreading support. A thread will be 
spawned for every successful HTTP connection. This thread will handle its own
associated connection and wait for a command before processing it and giving a 
reply. If a command is not received or completed after a specified amount of
time, a timeout will occur and the connection will be dropped along with its
corresponding thread being deallocated from the memory.

@@

Every time a modification or addition is done on a student's schedule, the 
backend will check for conflicts before executing the specified action. 
In order to implement this, the students schedule is stored as a set of numbers
that represent the index of each 15-minute block throughout the day beginning at
7:00 AM. For example, the number represents the time from 7:00AM to 7:15AM. This
representation is also applied to the schedule of each subject.

With the said representation of schedules, conflict checking is done by simply executing
a set intersection of the schedule of the student and the schedule of the subject. If the
resulting set does not have the size of zero, then it can be said that there is a conflict
between the two schedules.

@@

The system is designed to support HTTP Requests from specific domains
where the frontend code will be stored through the Access-Control-Allow-Origin header. 
This allows us to outsource the load added by frontend data being served as 
it will take a significant amount of networking bandwidth to serve the images and
scripts included. This will also allow the frontend to have heavier images and 
scripts without affecting the performance of the main server.

@@

The frontend is a static web application that fires REST requests to the
backend server. Requests use HTTP methods such as GET, POST, PUT, and DELETE to 
specify how data on the server is specified. Some of the commands have the same URL
but is executed by specifying the HTTP method. For example, a PUT request on the schedule
URL enlists a subject while a DELETE request on the same URL will cancel a subject. the
server will then respond with the result in Javascript Object Notation (JSON) format that
includes the resulting data and the return code specifying the kind of error or if there is
no error.

A copy of the list of subjects and schedules in JSON format is also loaded with the backend
making subject searching fast and reducing API server load by not sending requests that are known 
to fail such as enlistment and waitlisting of already-conflicting subjects.

The static file server of the frontend can be served on a server with a caching service
like CloudFlare enabled which would cache the static web application into its own 
Content Delivery Network(CDN).

Moreover, the static web application can also be cached on the browser of the client by
specifying Cache-Control headers. We can go even further by utilizing IndexDB, AppCache,
and localStorage capabilities to store the frontend application which will completely store
the application in the browser making it accessible even if the static server is down.

This setup also lessens the number of connections that the server will receive lessening
the number of threads running in the backend. This also alleviates the additional load 
caused by students opening the site in multiple tabs and refreshing each tab repeatedly.

The design for the frontend tries to follow Google's Material Design 
Specification although not strictly. Along with this, it shall support mobile
platforms by implementing Responsive Design.

@@

The system is evaluated by its stability and efficiency.

The system's stability and efficiency were evaluated by running benchmarks on the system.
Information such as requests per second, time per request, and transfer rate
of each function are collected. The system was also be tested if it can 
handle medium and heavy workloads by using load testing tools and services.
These tests will use combinations of the different functions at ratios expected
during actual use of the system.

Benchmarks are run on both the API server and the database management system.

The API server is benchmarked by running the tools wrk and apachebench. These two tools apply 
load to the server by running thousands of connections and recording either the number
of requests done after a certain duration or the time it takes for a specified number of requests to finish.

On the other hand, the database management system is benchmarked by running its own benchmarking tool
that can benchmark server-side scripts and simulate its interaction with a specified number of keys which
makes the evaluation of the speed of a script when there are a certain number of data on the server much easier.

\section{Timeline}
\begin{center}
    \begin{tabular}[!htbp]{ | l | p{5cm} |}
    \hline
      Date & Activity \\ \hline
      Day 0-7 & Planning of System Framework and Database Structures \\ \hline
      Day 8-14 & Implementation of Basic Backend Functions Without Conflict Checking \\ \hline
      Day 15-30 & Implementation of Conflict Checking and Integration with Backend \\ \hline
      Day 31-45 & Development of Static Web Application Interface \\ \hline
      Day 46-60 & Implementation of Advanced Features \\ \hline
      Day 61-68 & Stress Testing and Fuzzing \\ \hline
      Day 69-76 & Adjustments from the Results of Stress Testing and Fuzzing \\ \hline
      Day 77-81 & Alpha Test of Test Users \\ \hline
      Day 82-96 & Adjustments from the Results of Alpha Test \\ \hline
      Day 97-103 & Beta Release \\ \hline
    \end{tabular}
\end{center}

\section{Results and Discussion}
\begin{tikzpicture}
  \begin{axis}[ ybar, 
		legend style={at={(0.5,-0.2)}, 
		anchor=north,legend columns=-1}, 
		ylabel={1000x Req/s}, 
		symbolic x coords={Enlist, Cancel, Get Sched, Get Slots, Waitlist, Unwaitlist},
		xtick=data, 
		nodes near coords, 
		nodes near coords align={vertical}, 
		x tick label style={rotate=45,anchor=east}, ]
		\addplot coordinates {(Enlist,65.5) (Cancel,49.0) (Get Sched,140.9) (Get Slots,102.3) (Waitlist,69.0) (Unwaitlist,73.7)};
  \end{axis}
\end{tikzpicture}
\begin{center}
  Figure 1. Database requests per second
\end{center}
\begin{tikzpicture}
  \begin{axis}[ ybar, 
		legend style={at={(0.5,-0.2)}, 
		anchor=north,legend columns=-1}, 
		ylabel={1000x Req/s}, 
		symbolic x coords={Enlist, Cancel, Get Sched, Get Slots, Waitlist, Unwaitlist},
		xtick=data, 
		nodes near coords, 
		nodes near coords align={vertical}, 
		x tick label style={rotate=45,anchor=east}, ]
		\addplot coordinates {(Enlist,18.1) (Cancel,17.7) (Get Sched,17.1) (Get Slots,17.6) (Waitlist,17.5) (Unwaitlist,17.7)};
  \end{axis}
\end{tikzpicture}
\begin{center}
  Figure 2. API server requests per second on a Dual Core i7-5600U
\end{center}

\section{Summary and Conclusion}

% REFERENCES FORMAT
% References must be the same font size as other body text.
\bibliographystyle{sigchi}
\bibliography{sp2-paper}
